A sample workflow

### 2025-04-30 analyses/lpiworkflow.qmd
small change

What is global trend over time in vertebrate species population numbers, both overall and how does it vary acrosss species and populations?

@@

lpi_subset <- read.table(file.path(wd, "jdavies/African_mammals.txt"))

lpi_subset <- lpi_subset[, c("ID", "Genus", "Species", "Units", paste0("X", 1950:2020))]
lpi_subset$genus_species <- paste(lpi_subset$Genus, lpi_subset$Species)

lpi_subset_rsh <- 
  reshape(lpi_subset, direction = "long", idvar = "ID",
          varying =  paste0("X", 1950:2020), v.names = c("value"), timevar = "year", times = 1950:2020)
lpi_subset_rsh <- lpi_subset_rsh[lpi_subset_rsh$value != "NULL",]

@@

@@

First, we start with the parameters, which include an overall estimate of the trend across species (this parameter would address our question that we decided on above):

@@

lpi_subset <- read.table("/home/victor/projects/forecastflows/analyses/jdavies/African_mammals.txt")

@@

At this point, we would have a lot of ideas of ways to improve this model that we would want to follow up on, including (but not limited to) questioning our assumption of one sigma for all populations (CHECK HERE: do we really use one sigma for all populations?), how we model population trends (e.g., there are many ways to model abundance) and (relatedly) our linearity assumption. One possibility would be to move to a non-linear regression, and use a more flexible model that would allow population numbers to stabilize, such as shown using a logistic model for one wildbeest population in Tanzania:

>>>>

What is global trend over time in vertebrate species population numbers, both overall and how does it vary across species and populations?

@@

lpi_data <- read.csv(file.path(wd, "input", "LPD_2024_public.csv"), header=T)

@@

```{r}

lpi_subset <-subset(lpi_data, Region == "Africa" & Class == "Mammalia" & Included.in.LPR2024 == 1 )

# subset 5+ counts
obs.cnt<-NULL
for (n in 1:dim(lpi_subset)[1]){
  obs.cnt[n]<-sum(lpi_subset[n, 31:101]!="NULL")
}
lpi_subset<-lpi_subset[obs.cnt>4,]

# subset species 10+ populations
sp <- unique(lpi_subset$Binomial)
sp.cnt<-NULL
for(x in 1:length(sp)){
  Af.tmp<-subset(lpi_subset, Binomial == sp[x])
  sp.cnt[x]<-dim(Af.tmp)[1]
}
keep.sp<-sp[sp.cnt>9]
lpi_subset<-lpi_subset[lpi_subset$Binomial %in% keep.sp,]

@@

First, we start with the parameters, which include an overall estimate of the trend across species (this parameter would address our question that we decided on above). The red dashed line represents the true simulated value:

@@

@@

At this point, we would have a lot of ideas of ways to improve this model that we would want to follow up on, including (but not limited to) questioning our assumption of one observational error $\sigma$ for all populations, how we model population trends (e.g., there are many ways to model abundance) and (relatedly) our linearity assumption. One possibility would be to move to a non-linear regression, and use a more flexible model that would allow population numbers to stabilize, such as shown using a logistic model for one wildbeest population in Tanzania:
